Data processing

### 2024-04-20 manuscript.qmd
-

```{r}
library ("ggplot2") 
library ("gridExtra")
library ("dplyr")       
library ("paletteer")       
library ("tibble") # rownames_to_column
library ("tidyr") # pivot_longer
library ("ALFAM2")
library ("randomForest")
library ("xgboost")
library ("glmnet")
library ("shapviz")
library ("treeshap")

set.seed (123)
```

@@

```{r}
load (file = "scripts/data_alfam2.Rdata")
```

```{r}
plot = data_alfam2 %>%

@@

png (file = "figures/data_description_relative_cumul.png", width = 1000, height = 700)
plot
dev.off()

plot
```

@@

```{r}
pars <- alfam2pars01[!grepl('man.ph', names(alfam2pars01))]

@@

        # Important to use alfam2pars01 parameters, which correspond to the parameters used in the paper
        # New parameters are available, but leads to ohter results (less good results)
        pars = pars,

@@

```{r}
load (file = "scripts/data_random_forest.Rdata")
```

@@

random_forest_model = randomForest (

@@

        mtry = 22, nodesize = 3, replace = FALSE, sample_frac = 0.8

@@

```{r}
load (file = "scripts/data_xgboost.Rdata")
```

```{r}

xgboost_model = xgboost (

@@

    max.depth = 4, nrounds = 150, eta = 0.3, min_child_weight = 0.3, subsample = 1,

@@

```{r}
load (file = "scripts/data_lasso.Rdata")
```

```{r}
lasso_model = cv.glmnet (

@@

```{r}
plot = predictions_of_all_methods %>%

@@

png (file = "figures/observed_vs_predicted_values.png", width = 1000, height = 700)
plot
dev.off()

plot
```

## Evaluation metrics

@@

```{r}
# Evaluation for 72h cum. emission --------
df_plot = df_evaluation_metrics %>%
    filter (name == "72h cum. emission" & dataset == "Evaluation subset") %>%
    select (- name) %>%
    pivot_longer (cols = c (Pearsons_r, ME, MAE, MBE)) %>%
    mutate (name = factor (name, levels = c("Pearsons_r", "ME", "MAE", "MBE")))

suppressWarnings(ggplot (df_plot) +
    geom_histogram (aes (x = name, y = value, fill = method), position = "dodge", stat = "identity") +
    facet_wrap (~ name, scales = "free", nrow = 1) + 
    xlab ("") + ylab ("") + labs (fill = "") +
    theme (legend.position = "none ", strip.text.y = element_text(angle = 0), axis.text.x = element_blank(), axis.ticks.x = element_blank()) +
    scale_fill_manual (values = Dark2[c(2, 5, 6, 8)]) +
    ggtitle ("Evaluation subset") + theme (plot.title = element_text (margin= ggplot2::margin(0,0,20,0), hjust = 0, size = 30)))
# --------

# Evaluation for 72h relative cum. emission ----------
df_plot = df_evaluation_metrics %>%
    filter (name == "72h cum. emission" & dataset == "Calibration subset") %>%
    select (- name) %>%
    pivot_longer (cols = c (Pearsons_r, ME, MAE, MBE)) %>%
    mutate (name = factor (name, levels = c("Pearsons_r", "ME", "MAE", "MBE")))

suppressWarnings(ggplot (df_plot) +
    geom_histogram (aes (x = name, y = value, fill = method), position = "dodge", stat = "identity") +
    facet_wrap (~ name, scales = "free", nrow = 1) + 
    xlab ("") + ylab ("") + labs (fill = "") +
    theme (legend.position = "bottom", strip.text.y = element_text(angle = 0), axis.text.x = element_blank(), axis.ticks.x = element_blank()) +
    scale_fill_manual (values = Dark2[c(2, 5, 6, 8)]) +
    ggtitle ("Calibration subset") + theme (plot.title = element_text (margin= ggplot2::margin(0,0,20,0), hjust = 0, size = 30)))
# --------
```

@@

```{r}
random_forest_model = randomForest (

@@

```{r}
unified_model_rf <- randomForest.unify(random_forest_model, data_random_forest_calibration %>% select (- e.cum))

treeshap_rf <- treeshap(unified_model_rf, data_random_forest_calibration %>% select (- e.cum))

shapley_rf <- shapviz(treeshap_rf, X = data_random_forest_calibration %>% select (- e.cum))
```

```{r}
plot_shap_rf = sv_importance(shapley_rf, kind = "beeswarm", max_display = 10L) +
    scale_colour_gradient(low = "darkblue", high = "red", breaks = c(0, 1), labels = c("low", "high")) +
    theme (
      legend.title = element_text (size = 25, margin = ggplot2::margin (r = 50, b = 100)),
      legend.position = "bottom",
      legend.key.width = unit(3, "cm"),
      axis.title.x = element_text (margin = ggplot2::margin( b = 30))
    )

@@

```{r}
df_importance_rf = importance (random_forest_model) %>%
    as.data.frame %>%
    rownames_to_column (var = "Variable") %>%
    mutate (`%IncMSE` = `%IncMSE` / max (`%IncMSE`)) %>%
    mutate (type = "A")

@@

```{r}
xgboost_model = xgboost (

@@

unified_model_xgboost <- xgboost.unify(xgboost_model, as.matrix(df_tmp))

treeshap_xgboost <- treeshap(unified_model_xgboost, df_tmp)

shapley_xgb <- shapviz(treeshap_xgboost, X = df_tmp)
```

```{r}
plot_shap_xgboost = sv_importance(shapley_xgb, kind = "beeswarm", max_display = 10L) +
  scale_colour_gradient(low = "darkblue", high = "red", breaks = c(0, 1), labels = c("low", "high")) +
  theme (legend.title = element_text (size = 30, margin = ggplot2::margin (r = 50, b = 100)),
         legend.position = "bottom",
         legend.key.width = unit(3, "cm"),
         axis.title.x = element_text (margin = ggplot2::margin( b = 30)))

@@

```{r}
df_importance_xgb = xgb.importance (model = xgboost_model) %>%
    mutate (Gain = Gain / max (Gain)) %>%
    mutate (type = "B")

@@

plot_importance_xgboost
```
# Part 3 - Testing scenarios

```{r}
load (file = "scripts/data_scenarios.Rdata")

@@

```{r}
load (file = "scripts/data_scenarios_random_forest.Rdata")

@@

```{r}
load (file = "scripts/data_scenarios_xgboost.Rdata")

@@

    mutate (e.cum_hat_xgboost = xgboost_predictions_scenarios_vector, .before = time) %>%
    mutate (e.cum_hat_xgboost = ifelse (e.cum_hat_xgboost < 0, 0, e.cum_hat_xgboost)) %>%

@@

```{r}
load (file = "scripts/data_scenarios_lasso.Rdata")

@@

```{r}
load (file = "scripts/data_scenarios_alfam2.Rdata")

@@

```{r}
plot = data_scenarios %>% 
    select (app.mthd, incorp, t.incorp, man.ph,
            efficacy_random_forest, efficacy_xgboost, efficacy_lasso, efficacy_alfam2) %>%

@@

png (file = "figures/emissions_for_scenarios.png", width = 1000, height = 800)
plot
dev.off()

plot
```

@@

plot_shap_rf_2 = plot_shap_rf +
    annotate (geom = "label", label = "A", size = 22, fontface = "bold", x = 58, y = 1.5, fill = "black", color = "white") +
    xlim (c (- 20, 58)) +
    theme (legend.position = "none", 
          plot.margin = unit (c (0, 0, 1, 0), "cm"),
          axis.text.x = element_blank(),
          axis.title.x = element_blank(),
          axis.ticks.x = element_blank())

plot_shap_xgboost_2 = plot_shap_xgboost +
    annotate (geom = "label", label = "B", size = 22, fontface = "bold", x = 58, y = 1.5, fill = "black", color = "white") +
    xlim (c (- 20, 58)) + 
    theme (legend.position = "none", 
          plot.margin = unit (c (0, 0, 1, 0), "cm"),
          axis.title.x = element_text (margin = ggplot2::margin (t = 10, b = 20)))

>>>>

```{r}
library ("ggplot2") # version 3.4.4
library ("gridExtra") # version 2.3
library ("dplyr") # version 1.1.2      
library ("paletteer") # version 1.4.0
library ("tibble") # version 3.2.1
library ("tidyr") # version 1.3.0
library ("ALFAM2") # version 3.7
library ("randomForest") # version 4.7.1.1
library ("xgboost") # version 1.7.6.1
library ("glmnet") # version 4.1.8
library ("shapviz") # version 0.9.3
library ("treeshap") # version 0.3.1
```

@@

```{r}
load (file = "scripts/processed_data/data_alfam2.Rdata")
```

```{r, fig.width = 10, fig.height = 8}
plot_data_description = data_alfam2 %>%

@@

plot_data_description
```

```{r, include = FALSE}
png (file = "figures/data_description_relative_cumul.png", width = 1000, height = 700)
plot_data_description
dev.off()

```

```{r}
data_alfam2 %>%

    select (air.temp, wind.2m, rain.rate, rain.cum) %>%

    summary 
```

```{r}
data_alfam2 %>%

    filter (time == max (time), .by = pmid) %>%

    select (e.cum, time, tan.app, app.rate, man.dm, man.ph, t.incorp) %>%

    summary 
```

@@

```{r}
# We use alfam2pars01 parameter without pH, like in Hafner et al, 2019
pars <- alfam2pars01[!grepl('man.ph', names(alfam2pars01))]

@@

        pars = pars,

@@

```{r}
load (file = "scripts/processed_data/data_random_forest.Rdata")
```

@@

set.seed (123)
random_forest_model = randomForest (

@@

        mtry = 19, nodesize = 3, replace = FALSE, sample_frac = 0.8

@@

```{r}
load (file = "scripts/processed_data/data_xgboost.Rdata")
```

```{r}
set.seed (123)
xgboost_model = xgboost (

@@

    max.depth = 6, nrounds = 300, eta = 0.3, min_child_weight = 0.5, subsample = 0.8,

@@

```{r}
load (file = "scripts/processed_data/data_lasso.Rdata")
```

```{r}
set.seed (123)
lasso_model = cv.glmnet (

@@

```{r, fig.width = 10, fig.height = 8}
plot_observed_vs_predicted_values = predictions_of_all_methods %>%

@@

plot_observed_vs_predicted_values
```

```{r, include = FALSE}
png (file = "figures/observed_vs_predicted_values.png", width = 1000, height = 700)
plot_observed_vs_predicted_values
dev.off()
```

### Evaluation metrics

@@

```{r, fig.width = 16, fig.height = 5}
# Evaluation for 72h cum. emission
df_plot = df_evaluation_metrics %>%
    filter (name == "72h cum. emission" & dataset == "Evaluation subset") %>%
    select (- name) %>%
    pivot_longer (cols = c (Pearsons_r, ME, MAE, MBE)) %>%
    mutate (name = factor (name, levels = c("Pearsons_r", "ME", "MAE", "MBE")))

suppressWarnings(ggplot (df_plot) +
    geom_histogram (aes (x = name, y = value, fill = method), position = "dodge", stat = "identity") +
    facet_wrap (~ name, scales = "free", nrow = 1) + 
    xlab ("") + ylab ("") + labs (fill = "") +
    theme (legend.position = "none ", strip.text.y = element_text(angle = 0), axis.text.x = element_blank(), axis.ticks.x = element_blank()) +
    scale_fill_manual (values = Dark2[c(2, 5, 6, 8)]) +
    ggtitle ("Evaluation subset") + theme (plot.title = element_text (margin= ggplot2::margin(0,0,20,0), hjust = 0, size = 30)))
```

```{r, fig.width = 16, fig.height = 5}
# Evaluation for 72h relative cum. emission
df_plot = df_evaluation_metrics %>%
    filter (name == "72h cum. emission" & dataset == "Calibration subset") %>%
    select (- name) %>%
    pivot_longer (cols = c (Pearsons_r, ME, MAE, MBE)) %>%
    mutate (name = factor (name, levels = c("Pearsons_r", "ME", "MAE", "MBE")))

suppressWarnings(ggplot (df_plot) +
    geom_histogram (aes (x = name, y = value, fill = method), position = "dodge", stat = "identity") +
    facet_wrap (~ name, scales = "free", nrow = 1) + 
    xlab ("") + ylab ("") + labs (fill = "") +
    theme (legend.position = "bottom", strip.text.y = element_text(angle = 0), axis.text.x = element_blank(), axis.ticks.x = element_blank()) +
    scale_fill_manual (values = Dark2[c(2, 5, 6, 8)]) +
    ggtitle ("Calibration subset") + theme (plot.title = element_text (margin= ggplot2::margin(0,0,20,0), hjust = 0, size = 30)))
```

@@

```{r}
set.seed (123)
random_forest_model = randomForest (

@@

```{r}
df_tmp = data_random_forest_calibration %>% select (- e.cum)

unified_model_rf <- randomForest.unify (random_forest_model, df_tmp)

treeshap_rf <- treeshap (unified_model_rf, df_tmp, verbose = FALSE)

shapley_rf <- shapviz (treeshap_rf, X = data_random_forest_calibration %>% select (- e.cum))
```

```{r, fig.width = 10, fig.height = 8}
plot_shap_rf = sv_importance(shapley_rf, kind = "beeswarm", max_display = 10L) +
    scale_colour_gradient(low = "darkblue", high = "red", breaks = c(0, 1), labels = c("low", "high")) +
    theme (
      legend.title = element_text (size = 25, margin = ggplot2::margin (r = 50, b = 100)),
      legend.position = "bottom",
      legend.key.width = unit(3, "cm"),
      axis.title.x = element_text (margin = ggplot2::margin( b = 30))
    )

@@

```{r, fig.width = 10, fig.height = 12}
df_importance_rf = importance (random_forest_model) %>%
    as.data.frame %>%
    rownames_to_column (var = "Variable") %>%
    mutate (`%IncMSE` = `%IncMSE` / max (`%IncMSE`)) %>%
    mutate (type = "A")

@@

```{r}
set.seed (123)
xgboost_model = xgboost (

@@

unified_model_xgboost <- xgboost.unify (xgboost_model, as.matrix(df_tmp))

treeshap_xgboost <- treeshap (unified_model_xgboost, df_tmp, verbose = FALSE)

shapley_xgb <- shapviz (treeshap_xgboost, X = df_tmp)
```

```{r, fig.width = 10, fig.height = 8}
plot_shap_xgboost = sv_importance(shapley_xgb, kind = "beeswarm", max_display = 10L) +
  scale_colour_gradient(low = "darkblue", high = "red", breaks = c(0, 1), labels = c("low", "high")) +
  theme (legend.title = element_text (size = 30, margin = ggplot2::margin (r = 50, b = 100)),
         legend.position = "bottom",
         legend.key.width = unit(3, "cm"),
         axis.title.x = element_text (margin = ggplot2::margin( b = 30)))

@@

```{r, fig.width = 10, fig.height = 12}
df_importance_xgb = xgb.importance (model = xgboost_model) %>%
    mutate (Gain = Gain / max (Gain)) %>%
    mutate (type = "B")

@@

plot_importance_xgboost
```

```{r}
set.seed (123)
lasso_model = cv.glmnet (

    x = data_lasso %>% 
        select (- pmid, - e.cum, - e.rel, - dataset) %>%
        as.matrix %>%
        {.},

    y = data_lasso %>% 
        select (e.cum) %>% 
        as.matrix %>%
        {.},

    alpha = 1

# Part 3 - Testing scenarios

```{r}
load (file = "scripts/processed_data/data_scenarios.Rdata")

@@

```{r}
load (file = "scripts/processed_data/data_scenarios_random_forest.Rdata")

@@

```{r}
load (file = "scripts/processed_data/data_scenarios_xgboost.Rdata")

@@

    mutate (e.cum_hat_xgboost = xgboost_predictions_scenarios_vector, .before = time) %>%

@@

```{r}
load (file = "scripts/processed_data/data_scenarios_lasso.Rdata")

@@

```{r}
load (file = "scripts/processed_data/data_scenarios_alfam2.Rdata")

@@

```{r, fig.width = 10, fig.height = 10}
plot_scenario_predictions = data_scenarios %>% 
    select (app.mthd, incorp, t.incorp, man.ph,
            efficacy_random_forest, efficacy_xgboost, efficacy_lasso, efficacy_alfam2) %>%

@@

plot_scenario_predictions
```

```{r, include = FALSE}
png (file = "figures/emissions_for_scenarios.png", width = 1000, height = 800)
plot_scenario_predictions
dev.off()
```

@@

plot_shap_rf_2 = plot_shap_rf +
    annotate (geom = "label", label = "A", size = 22, fontface = "bold", x = 45, y = 1.5, fill = "black", color = "white") +
    xlim (c (- 18, 45)) +
    theme (legend.position = "none", 
          plot.margin = unit (c (0, 0, 1, 0), "cm"),
          axis.text.x = element_blank(),
          axis.title.x = element_blank(),
          axis.ticks.x = element_blank())

plot_shap_xgboost_2 = plot_shap_xgboost +
    annotate (geom = "label", label = "B", size = 22, fontface = "bold", x = 45, y = 1.5, fill = "black", color = "white") +
    xlim (c (- 18, 45)) + 
    theme (legend.position = "none", 
          plot.margin = unit (c (0, 0, 1, 0), "cm"),
          axis.title.x = element_text (margin = ggplot2::margin (t = 10, b = 20)))

### 2024-04-20 scripts/data_processing.qmd
-

```{r}
library ("ggplot2") 
library ("dplyr")       
library ("paletteer")       
library ("tibble") # rownames_to_column
library ("tidyr") # pivot_longer
```

```{r}
axis_text_size = 24 
axis_title_size = 30
title_size = 26

@@

```{r}
data_interval_level = read.csv (file = "../data/ALFAM2_interval.csv")
data_plot_level = read.csv (file = "../data/ALFAM2_plot.csv")

calibration_subset_pmids = read.csv (file = "../data/1-s2.0-S1352231018308069-mmc1.csv")
evaluation_subset_pmids = read.csv (file = "../data/1-s2.0-S1352231018308069-mmc2.csv")

@@

    filter (ct > 0 & ct <= 78) %>%
    select (j.NH3, e.cum, e.rel, pmid, ct, dt, air.temp, wind.2m, rain.rate, rain.cum, dataset)  %>%

@@

    # we keep only trials that last more than 24h, to be in the same settings as for the machine learning models
    filter (pmid %in% pmid [max (ct) > 24.5], .by = pmid) %>%

    rename (time = ct) %>%

save (data_alfam2, file = "data_alfam2.Rdata")
```

@@

    # we round dt and time in order to compute the new climatic variables
    # cf the next chunk
    mutate (dt = round (dt, digits = 1)) %>%
    mutate (time = cumsum (dt), .by = pmid) %>%

@@

```{r}
# data = data_ml
# pmids = data %>% unique_vector (pmid)
# 
# data_ml = NULL
# 
# for (i in pmids) {
#         
#     data_tmp = data %>% filter (pmid == i)
#     
#     time_max = max (data_tmp$time)
#     
#     time = seq (0, (time_max - 0.1), by = 0.1)
#     
#     group = c(
#         sapply (c (1 : 6), function (j) rep (j, 40)) %>% as.vector,
#         rep (7, (length (time) - 240))
#     )
#     
#     # here we added 1+e-8 for numerical reasons, ('19.8 - 16.7 <= 3.1' return FALSE)
#     temp = sapply (time, function (j) data_tmp %>% filter ((time - dt) <= j + 1e-8 & j + 1e-8 < time) %>% pull (air.temp))
#     wind = sapply (time, function (j) data_tmp %>% filter ((time - dt) <= j + 1e-8 & j + 1e-8 < time) %>% pull (wind.2m))
#     rain = sapply (time, function (j) data_tmp %>% filter ((time - dt) <= j + 1e-8 & j + 1e-8 < time) %>% pull (rain.rate))
# 
# 
#     df_temp = data.frame (time = time, temp = temp, group = group) %>% 
# 
#         summarise (temp = mean (temp), .by = group) %>% 
#         pivot_wider (names_from = group, values_from = temp, names_prefix = "temp_")
#                    
# 
#     df_wind = data.frame (time = time, wind = wind, group = group) %>% 
# 
#         summarise (wind = mean (wind), .by = group) %>% 
#         pivot_wider (names_from = group, values_from = wind, names_prefix = "wind_")
#                    
# 
#     df_rain = data.frame (time = time, rain = rain, group = group) %>% 
# 
#         summarise (rain = mean (rain), .by = group) %>% 
#         pivot_wider (names_from = group, values_from = rain, names_prefix = "rain_")
# 
# 
#     data_tmp = cbind (
#         data_tmp[nrow (data_tmp), ] %>% select (dataset, e.cum, e.rel, pmid, time, tan.app, app.mthd, app.rate, man.dm, man.ph, man.source, incorp, t.incorp),
#         df_temp,
#         df_wind,
#         df_rain
#     )
#                    
#     data_ml = rbind (data_ml, data_tmp)
#                                                          
# }

# save (data_ml, file = "data_ml.Rdata")

load (file = "data_ml.Rdata")
```

@@

save (data_random_forest, file = "data_random_forest.Rdata")
```

@@

save (data_xgboost, file = "data_xgboost.Rdata")
```

@@

save (data_lasso, file = "data_lasso.Rdata")
```

@@

```{r}
data_ml %>%

    select (rain_1, rain_2, rain_3, rain_4, rain_5, rain_6, rain_7) %>%

    apply (2, quantile, probs = seq (0.01, 1, by = 0.01)) %>%
    as.data.frame %>%
    mutate (quantile = row_number () / 100) %>%
    pivot_longer (cols = - c (quantile)) %>%

    filter (quantile > 0.75) %>%
    ggplot () +
        geom_line (aes (x = quantile, y = value, color = name)) +
        scale_color_manual (values = Dark2)

@@

    select (temp_1, temp_2, temp_3, temp_4, temp_5, temp_6, temp_7,
            wind_1, wind_2, wind_3, wind_4, wind_5, wind_6, wind_7) %>%

@@

    select (rain_1, rain_2, rain_3, rain_4, rain_5, rain_6, rain_7) %>%

@@

```{r}
grid_climatic_variable = quartiles_climatic_variable [, c(1 : 7)] %>% 
    cross_join (quartiles_climatic_variable [, c(8 : 14)]) %>% 
    cross_join (quartiles_climatic_variable [, c(15 : 21)]) %>%

    mutate (group_temp = ifelse (temp_1 == quartiles_climatic_variable[1, 1], "q1", "q2")) %>%
    mutate (group_wind = ifelse (wind_1 == quartiles_climatic_variable[1, 8], "q1", "q2")) %>%
    mutate (group_rain = ifelse (rain_2 == quartiles_climatic_variable[1, 16], "q1", "q2"))
```

@@

@@

    # reordering the variables
    select (pmid, time, tan.app, app.mthd, app.rate, man.dm, man.ph, man.source, incorp, t.incorp,
            temp_1, temp_2, temp_3, temp_4, temp_5, temp_6, temp_7,
            wind_1, wind_2, wind_3, wind_4, wind_5, wind_6, wind_7,
            rain_1, rain_2, rain_3, rain_4, rain_5, rain_6, rain_7,
            group_temp, group_wind, group_rain)

save (data_scenarios, file = "data_scenarios.Rdata")
```

@@

save (data_scenarios_random_forest, file = "data_scenarios_random_forest.Rdata")
```

@@

save (data_scenarios_xgboost, file = "data_scenarios_xgboost.Rdata")
```

@@

save (data_scenarios_lasso, file = "data_scenarios_lasso.Rdata")
```

@@

    pivot_longer (cols = c (10 : 30), names_to = c (".value", "time"), names_pattern = "(.+)_(.+)") %>%

@@

    mutate (time = as.numeric (recode (time, "1" = "4", "2" = "8", "3" = "12", "4" = "16", "5" = "20", "6" = "24", "7" = "72"))) %>%
    mutate (dt = ifelse (time == 72, (72 - 24), 4), .after = time) %>%

@@

save (data_scenarios_alfam2, file = "data_scenarios_alfam2.Rdata")
```

>>>>

In this file, we build several dataframes that are used in the manuscript.qmd file. The following figure summarizes how these dataframes are built.

![](figures/code_explanations.png){width=500, fig-align="center"}

From the files of the ALFAM2 database (available in a Zenodo repository at https://zenodo.org/records/8109634), we first construct data_alfam2, which corresponds to the dataset used in Hafner et al (2019). From data_alfam2, we construct data_ml by retaining only the last observation of each PMID. The dynamics of climatic variables are preserved through class variables. From data_ml, we create data_lasso, data_xgboost, and data_random_forest, all of which contain the same information but are formatted according to the packages used to build the respective models. Additionally, from data_ml, we generate data_scenarios, which correspond to various scenarios on which we test the models. These scenarios are based on quantiles derived from data_ml. From data_scenarios we create four dataframes containing the same information but formatted according to the packages used for the different models.

```{r}
library ("ggplot2") 
library ("dplyr")       
library ("paletteer")       
library ("tidyr")
```

```{r, include = FALSE}
axis_text_size = 24 
axis_title_size = 30
title_size = 26

@@

```{r}
# these two files come from the ALFAM2 Zenodo repository (https://zenodo.org/records/8109634)
data_interval_level = read.csv (file = "../data/ALFAM2_interval.csv")
data_plot_level = read.csv (file = "../data/ALFAM2_plot.csv")

# these two files come from the supplementary material of Hafner et al (2019)
calibration_subset_pmids = read.csv (file = "../data/1-s2.0-S1352231018308069-mmc1.csv")
evaluation_subset_pmids = read.csv (file = "../data/1-s2.0-S1352231018308069-mmc2.csv")

@@

    filter (ct <= 78) %>%
    select (j.NH3, e.cum, e.rel, pmid, ct, dt, air.temp, wind.2m, rain.rate, rain.cum, dataset)  %>%

@@

    rename (time = ct)

save (data_alfam2, file = "processed_data/data_alfam2.Rdata")
```

@@

    # we round dt and time in order to compute the new climatic variables
    # cf the next chunk
    mutate (dt = round (dt, digits = 2)) %>%
    mutate (time = cumsum (dt), .by = pmid) %>%

@@

```{r}
# data = data_ml
# pmids = data %>% pull (pmid) %>% unique
# 
# data_ml = NULL
# 
# for (i in pmids) {
#         
#     data_tmp = data %>% filter (pmid == i)
#     
#     time_max = max (data_tmp$time)
#     
#     time = seq (0, (time_max - 0.01), by = 0.01)
#     
#     group = c(
#         sapply (c (1 : 5), function (j) rep (j, 400)) %>% as.vector,
#         rep (6, (length (time) - 2000))
#     )
#     
#     # here we added 1+e-8 for numerical reasons, ('19.8 - 16.7 <= 3.1' return FALSE)
#     temp = sapply (time, function (j) data_tmp %>% filter ((time - dt) <= j + 1e-8 & j + 1e-8 < time) %>% pull (air.temp))
#     wind = sapply (time, function (j) data_tmp %>% filter ((time - dt) <= j + 1e-8 & j + 1e-8 < time) %>% pull (wind.2m))
#     rain = sapply (time, function (j) data_tmp %>% filter ((time - dt) <= j + 1e-8 & j + 1e-8 < time) %>% pull (rain.rate))
# 
# 
#     df_temp = data.frame (time = time, temp = temp, group = group) %>% 
# 
#         summarise (temp = mean (temp), .by = group) %>% 
#         pivot_wider (names_from = group, values_from = temp, names_prefix = "temp_")
#                    
# 
#     df_wind = data.frame (time = time, wind = wind, group = group) %>% 
# 
#         summarise (wind = mean (wind), .by = group) %>% 
#         pivot_wider (names_from = group, values_from = wind, names_prefix = "wind_")
#                    
# 
#     df_rain = data.frame (time = time, rain = rain, group = group) %>% 
# 
#         summarise (rain = mean (rain), .by = group) %>% 
#         pivot_wider (names_from = group, values_from = rain, names_prefix = "rain_")
# 
# 
#     data_tmp = cbind (
#         data_tmp[nrow (data_tmp), ] %>% select (dataset, e.cum, e.rel, pmid, time, tan.app, app.mthd, app.rate, man.dm, man.ph, man.source, incorp, t.incorp),
#         df_temp,
#         df_wind,
#         df_rain
#     )
#                    
#     data_ml = rbind (data_ml, data_tmp)
#                                                          
# }

# save (processed_data/data_ml, file = "data_ml.Rdata")

load (file = "processed_data/data_ml.Rdata")
```

@@

save (data_random_forest, file = "processed_data/data_random_forest.Rdata")
```

@@

save (data_xgboost, file = "processed_data/data_xgboost.Rdata")
```

@@

save (data_lasso, file = "processed_data/data_lasso.Rdata")
```

@@

```{r, fig.width = 10, fig.height = 5}
data_ml %>%

    select (rain_1, rain_2, rain_3, rain_4, rain_5, rain_6) %>%

    apply (2, quantile, probs = seq (0.01, 1, by = 0.01)) %>%
    as.data.frame %>%
    mutate (quantile = row_number () / 100) %>%
    pivot_longer (cols = - c (quantile), names_to = "variable") %>%

    filter (quantile > 0.75) %>%
    ggplot () +
        geom_line (aes (x = quantile, y = value, color = variable)) +
        scale_color_manual (values = Dark2)

@@

    select (temp_1, temp_2, temp_3, temp_4, temp_5, temp_6,
            wind_1, wind_2, wind_3, wind_4, wind_5, wind_6) %>%

@@

    select (rain_1, rain_2, rain_3, rain_4, rain_5, rain_6) %>%

@@

```{r}
quartiles_climatic_variable
```

```{r}
grid_climatic_variable = quartiles_climatic_variable [, c(1 : 6)] %>% 
    cross_join (quartiles_climatic_variable [, c(7 : 12)]) %>% 
    cross_join (quartiles_climatic_variable [, c(13 : 18)]) %>%

    mutate (group_temp = ifelse (temp_1 == quartiles_climatic_variable[1, 1], "q1", "q2")) %>%
    mutate (group_wind = ifelse (wind_1 == quartiles_climatic_variable[1, 7], "q1", "q2")) %>%
    mutate (group_rain = ifelse (rain_2 == quartiles_climatic_variable[1, 13], "q1", "q2"))
```

@@

```{r}
quartiles_plot_level_variable
```

@@

    # reordering the variables
    select (pmid, time, tan.app, app.mthd, app.rate, man.dm, man.ph, man.source, incorp, t.incorp,
            temp_1, temp_2, temp_3, temp_4, temp_5, temp_6,
            wind_1, wind_2, wind_3, wind_4, wind_5, wind_6,
            rain_1, rain_2, rain_3, rain_4, rain_5, rain_6,
            group_temp, group_wind, group_rain)

save (data_scenarios, file = "processed_data/data_scenarios.Rdata")
```

@@

save (data_scenarios_random_forest, file = "processed_data/data_scenarios_random_forest.Rdata")
```

@@

save (data_scenarios_xgboost, file = "processed_data/data_scenarios_xgboost.Rdata")
```

@@

save (data_scenarios_lasso, file = "processed_data/data_scenarios_lasso.Rdata")
```

@@

    pivot_longer (cols = c (10 : 27), names_to = c (".value", "time"), names_pattern = "(.+)_(.+)") %>%

@@

    mutate (time = as.numeric (recode (time, "1" = "4", "2" = "8", "3" = "12", "4" = "16", "5" = "20", "6" = "72"))) %>%
    mutate (dt = ifelse (time == 72, (72 - 20), 4), .after = time) %>%

@@

save (data_scenarios_alfam2, file = "processed_data/data_scenarios_alfam2.Rdata")
```
